---
title: "Vignette for package dinoR"
author: " Michaela Schwaiger"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{dinoR-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
suppressPackageStartupMessages({
  library(dinoR)
  library(ggplot2)
  library(dplyr)
  })
```

#### load the NomeMatrix data for 2 replicates Adnp_KO and WT

```{r load data}
NomeMatrix <- readRDS(file = file.path("/tungstenfs/scratch/gbuehler/bioinfo/Rpackages/data_preparation/",
                                       "Nome_matrix_GCHamp.rds"))
NomeMatrix
```

#### generate read counts for TF open nuc 

```{r read counts}
footprint_counts <- footprintQuant(NomeMatrix,nr=10)
#remove any rows where no reads could be assigned a footprint
footprint_counts <- footprint_counts[rowSums(footprint_counts[,3:7]) > 0,]
footprint_counts
```


#### calculate differential NOMe-seq footprint abundance between Adnp KO and WT 

```{r diNOMe}
res <- diNOMeTest(footprint_counts,WTsamples = c("WT_1","WT_2"),
  KOsamples = c("AdnpKO_1","AdnpKO_2"))
res
```
#### plot the number of regulated ROIs within each ROI type

```{r nregulated, fig.height=4,fig.width=7}
res$ampType <- matrix(unlist(strsplit(res$ROI,"_")),ncol=4,byrow=TRUE)[,1]

res %>% group_by(contrasts,ampType,regulated) %>% summarize(n=n()) %>% 
  ggplot(aes(x=ampType,y=n,fill=regulated)) + geom_bar(stat="identity") + 
  facet_grid(~contrasts) + theme_bw() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
  scale_fill_manual(values=c("orange","grey","blue3"))
```

#### MA plots

```{r MAplot,fig.height=2,fig.width=10}
ggplot(res,aes(y=logFC,x=logCPM,col=regulated)) + geom_point() + 
  facet_grid(~contrasts) + theme_bw() + 
  scale_color_manual(values=c("orange","grey","blue3"))
```

#### calculate the percentage of fragments in each footprint type and plot a heatmap comparing percentages in WT and KO

```{r percentages,fig.height=5,fig.width=7}
ROIgroup <-  matrix(unlist(strsplit(footprint_counts$ROI,"_")),ncol=4,byrow=TRUE)[,1]
footprint_percentages <- footprintPerc(footprint_counts,ROIgroup)

#plot heatmaps 
fpPercHeatmap(footprint_percentages)
```

#### compare the footprint percentages and significance testing results for Adnp KO and WT 

```{r comparison,fig.height=5,fig.width=9}

compareFootprints(footprint_percentages,res,WTsamples = c("WT_1","WT_2"),
                  KOsamples = c("AdnpKO_1","AdnpKO_2"),plotcols = c("#f03b20", "#a8ddb5", "#bdbdbd"))

```

